---
title: "Determining Fitness For Coaching Decision Making"
author: "Jack Robinson, Josh Ogilvie, Josh Pawliuk"
date: "2026-03-07"
format: html
editor: visual
bibliography: references.bib
nocite: |
  @*
embed-resources: true
---

## Summary bullets:

-   Fitness can be defined by the effect that a runners heart rate has on their pace

-   Both runners in this analysis have a similar fitness level based on the current data

-   Over time, both runners have improved their average pace, however, only Collie was able to improve their average heart rate over time, this indicates that Collie was able to improve more in their fitness than Hound

-   Outdoor seasonality analysis follows more of an expected distribution for Collie, where Hound seems to be faster in cooler months like February and March due to competing in a faster-paced race

-   Comparing the Runners Aerobic TE we find Colllie to be running at a much higher Aerobic TE than Hound, indicating higher effort

## What is Fitness

```{r}
#| echo: false
#| message: false

library(tidyr)
library(tidyverse)
library(stats)
library(broom)
library(gt)
library(tsibble)
library(feasts)
library(splines)
library(scales)
library(patchwork)
library(bibtex)
```

```{r}
#| echo: false
#| message: false

url <- "https://raw.githubusercontent.com/joshua-ogilvie/Determining-Fitness-For-Coaching-Decision-Making/main/references.bib"

download.file(url, destfile = "references.bib", mode = "w")

```

```{r,Load Data,}
#| echo: false
#| message: false
full_data <- RTLedu::strava
```

```{r, Dropping unnecessary column variables}
#| echo: false
#| message: false

run_data <- full_data %>%
  dplyr::select(-Favorite, 
                -Avg.Vertical.Ratio, 
                -Avg.Vertical.Oscillation, 
                -Avg.Ground.Contact.Time, 
                -Training.Stress.Score., 
                -Avg.Power, 
                -Max.Power, 
                -Flow, 
                -Avg..Swolf, 
                -Avg.Stroke.Rate, 
                -Dive.Time, 
                -Surface.Interval,
                -Decompression, 
                -Avg.Resp, 
                -Min.Resp, 
                -Max.Resp
                )
```

```{r, change all distances to Km with track running type if not just keep distance and all Time values that we use in analysis to seconds }
#| echo: false
#| message: false


#Converting the varying distances into KM
run_data <- run_data %>%
  dplyr::mutate(Distance.km = ifelse(Activity.Type == "Track Running",
                                     Distance / 1000,
                                     Distance))

#lubridate was failing to parse properly and failed to render, here we convert our (HMS) data into a character then separated the characters by the ":" deliminator then used the values to reach seconds for all time values. 

run_data$Time <- as.character(run_data$Time)
run_data$Moving.Time <- as.character(run_data$Moving.Time) 
run_data$Avg.Pace <- as.character(run_data$Avg.Pace) 
run_data$Avg.HR <- as.numeric(run_data$Avg.HR)

run_data <- run_data %>% 
   tidyr::separate(Avg.Pace, 
                   into = c("p_m", "p_s", "p_mil"), 
                   sep = ":", 
                   convert = TRUE, 
                   remove = FALSE)%>%
   dplyr::mutate(avg_pace_seconds = p_m * 60 + p_s + p_mil / 1000) %>% 
  
  tidyr::separate(Time, 
                  into = c("t_h", "t_m", "t_s"), 
                  sep = ":", 
                  convert = TRUE, 
                  remove = FALSE) %>%
   dplyr::mutate(time_seconds = t_h * 3600 + t_m * 60 + t_s) %>% 
  
  tidyr::separate(Moving.Time, 
                  into = c("mt_h", "mt_m", "mt_s"), 
                  sep = ":", 
                  convert = TRUE, 
                  remove = FALSE) %>%
   dplyr::mutate(moving_time_seconds = mt_h * 3600 + mt_m * 60 + mt_s) %>% 
  
  dplyr::mutate(time_measured = ifelse(moving_time_seconds ==0, time_seconds, moving_time_seconds)) 
  
```

To determine which Runner is fitter, we must first define fitness. It may be easy to look at some run times and distances and try to determine who the better runner is, but this does not help us understand which runner is fitter. A common and very robust measurement for determining fitness is a VO2 max test, which assesses “the body’s ability to take in, transport, and use oxygen during exercise” [@UofA]. This test is also costly and requires advanced instruments to achieve accurate results. In this analysis, we will be using an approach that allows us to analyze a runner's fitness using data from Strava. Fitness level in this analysis will be determined by the impact that a runner's heart rate has on their pace. This proxy for aerobic efficiency gives us an understanding of which runner has a more efficient exchange of oxygen, or who is fitter. Using the Strava data, the runners' fitness levels can be evaluated by examining the relationship between their running pace and heart rate. A larger improvement in pace for the same increase in heart rate would indicate a greater level of aerobic fitness. In other words, the runner whose pace improves more for a given increase in heart rate can translate physiological effort into performance more effectively. We can use this metric to determine both who is currently more fit and also how much increase in fitness each runner has achieved over the timespan of the available data. In Figure 1, a side-by-side graph shows the average pace in seconds per km of both Collie and Hound. In Figure 2, a similar graph with a depiction of their average heart rate over time. These observations will be the foundation for our analysis and will be used to determine who the fitter runner is. 

```{r, Histograms of Average Pace and Average Heart rate}
#| echo: false
#| message: false
#| fig-align: center
 

pace_data <- run_data %>% 
  dplyr::group_by(Runner) %>% 
  dplyr::select(avg_pace_seconds, Date, Runner) %>% 
  dplyr::mutate(mean.pace = mean(avg_pace_seconds), 
                sd.pace = sd(avg_pace_seconds)) %>% 
  
  ggplot2::ggplot(ggplot2::aes(x = Date, y = avg_pace_seconds, color = Runner)) +
  ggplot2::geom_line() +
  ggplot2::labs(y = "Average Pace (Seconds / KM)",
                title = "(Figure-1) Average Pace Over Time") + 
  ggplot2::facet_wrap(Runner~., scales = "free_x") +
  ggplot2::geom_errorbar(ggplot2::aes( ymin = mean.pace - sd.pace , ymax = mean.pace + sd.pace), 
                alpha = 0.2, 
                col = "yellow")+
  scale_colour_manual(values = c("Collie" = "darkgreen","Hound" = "purple"))


hr_data <- run_data %>% 
  dplyr::group_by(Runner) %>% 
  dplyr::select(Avg.HR, Date, Runner) %>% 
  dplyr::mutate(mean.hr = mean(Avg.HR), 
                sd.hr = sd(Avg.HR))%>% 
  
  ggplot2::ggplot(ggplot2::aes(x = Date, y = Avg.HR, color = Runner)) +
  ggplot2::geom_line() +
  ggplot2::labs(y = "Average HR",
                title = "(Figure-2) Average HR Over Time") + 
  ggplot2::facet_wrap(Runner~., scales = "free_x") +
  ggplot2::geom_errorbar(ggplot2::aes( ymin = mean.hr - sd.hr , ymax = mean.hr + sd.hr), 
                alpha = 0.2, 
                col = "yellow", 
                fill)+
   scale_colour_manual(values = c("Collie" = "darkgreen","Hound" = "purple"))

pace_data/hr_data
```

## Which Athlete Has Better Fitness

Looking at the graphical analysis from Figure 1 of pace and heart rate over time, it initially appears that Collie is fitter. Over time, they appear to generally maintain both a lower average pace and a lower average heart rate compared to Hound while also displaying less variability in both metrics. It is important to note that many observations from Hound contained a run that was adjacent to another workout or a short warmup before executing more effort at another activity. We decided to remove these observations from our analysis of their running fitness. However, when examining regression results as shown in Figure 2, where the average pace is modelled as a function of heart rate, both runners show very similar coefficients on heart rate, Collie from Figure 3 at -2.0431 and Hound from Figure 4 at -2.0915. This implies that for each one BPM increase in average heart rate, Collie improves their average pace by about 2.04 seconds, and Hound improves theirs by slightly better, 2.09 seconds. Because fitness is being defined as aerobic efficiency, that is, how much pace changes for a given change in heart rate, these nearly identical coefficients suggest that the runners have very similar levels of aerobic efficiency. While Hounds' slightly larger result may imply a marginally better responsiveness of pace to an increase in heart rate, the R-squared in Figure 4 is quite low, meaning that heart rate alone is not a significant predictor of pace, reflecting the influence of other factors. At the same time, the very low p-values suggest that these relationships between pace and heart rate are statistically significant, thus heart rate does have a measurable effect on pace.

```{r,graphs of regression with residuals}
#| echo: false
#| message: false
#| warning: false
#| fig-align: center

#Adding column of number of observations due to double observations in some days and removing runs that are titled as "Cool Down" or with an adjacent "AB" workout
reg_data <- run_data %>% 
  dplyr::filter(!grepl("Cool|AB", Title)) %>% 
  dplyr::mutate(obs = c(1:length(Date))) %>% 
  dplyr::mutate(speed = 1/avg_pace_seconds) 
#%>% 
  #dplyr::mutate(hr_ef = speed/Avg.HR)

regc_dat <- reg_data %>% 
  dplyr::select(obs, avg_pace_seconds, Avg.HR, Runner, Aerobic.TE) %>% 
  dplyr::filter(Runner == "Collie") %>% 
  tidyr::pivot_wider(names_from = Runner,
                     values_from = c(avg_pace_seconds, Avg.HR)) %>% 
  tidyr::drop_na()


regh_dat <- reg_data %>% 
  dplyr::select(obs, avg_pace_seconds, Avg.HR, Runner, Aerobic.TE) %>% 
  dplyr::filter(Runner == "Hound") %>% 
  tidyr::pivot_wider(names_from = Runner,
                     values_from = c(avg_pace_seconds, Avg.HR, )) %>% 
  tidyr::drop_na()



regc <- stats::lm(avg_pace_seconds_Collie ~  Avg.HR_Collie , 
                  data = regc_dat)
regh <- stats::lm(avg_pace_seconds_Hound ~  Avg.HR_Hound , 
                  data = regh_dat)

regression_c_augmented <- broom::augment(regc)
regression_c_plot <- regression_c_augmented %>%
  ggplot2::ggplot(ggplot2::aes(x = Avg.HR_Collie, y = avg_pace_seconds_Collie)) +
  ggplot2::geom_point() +
  ggplot2::geom_smooth(method = "lm", formula = y ~  x ,
                       se = FALSE, color = "darkgreen") +
  ggplot2::geom_segment(ggplot2::aes(x = Avg.HR_Collie, xend = Avg.HR_Collie, y = .fitted, yend = avg_pace_seconds_Collie, color = "yellow")) +
  ggplot2::scale_x_continuous(limits = c(130,180))+
  ggplot2::labs(
    title = "(Figure-3) Collie Avg HR vs Avg Pace Relationship",
    subtitle = "2020 to 2023",
    x = "Avg HR (BPM)",
    y = "Avg Pace (seconds)"
  ) +
  ggplot2::theme(legend.position = "none")

regression_h_augmented <- broom::augment(regh)
regression_h_plot <- regression_h_augmented %>%
  ggplot2::ggplot(ggplot2::aes(x = Avg.HR_Hound, y = avg_pace_seconds_Hound)) +
  ggplot2::geom_point() +
  ggplot2::geom_smooth(method = "lm", se = FALSE, color = "purple") +
  ggplot2::geom_segment(ggplot2::aes(x = Avg.HR_Hound, xend = Avg.HR_Hound, y = .fitted, yend = avg_pace_seconds_Hound, color = "yellow")) +
  ggplot2::scale_x_continuous(limits = c(130,180))+
  ggplot2::labs(
    title = "(Figure-4) Hound Avg HR vs Avg Pace Relationship",
    subtitle = "2018 to 2023",
    x = "Avg HR (BPM)",
    y = "Avg Pace (seconds)"
  ) +
  ggplot2::theme(legend.position = "none")

regression_c_plot/ regression_h_plot

```

```{r, Regression}
#| echo: false
#| message: false 
#| fig-align: center

regc_summary_model <- summary(regc) %>%
  glance() %>%
  dplyr::select(-statistic,
                -df,
                -df.residual,
                -nobs) %>%
  dplyr::mutate(" " = "Model") %>%
  dplyr::rename("Adj R-Squared" = "adj.r.squared",
                "Residual SE" = "sigma",
                "P-value" = "p.value")

regh_summary_model <- summary(regh) %>%
  glance() %>%
  dplyr::select(-statistic,
                -df,
                -df.residual,
                -nobs) %>%
  dplyr::mutate(" " = "Model") %>%
  dplyr::rename("Adj R-Squared" = "adj.r.squared",
                "Residual SE" = "sigma",
                "P-value" = "p.value")

regc_model_table <- regc_summary_model %>%
  gt(rowname_col = " ",
     auto_align = FALSE) %>%
  tab_header(title = "(Figure-5) Model Statistics - Collie", 
             subtitle = "Regression of Average Pace on Heart Rate") %>%
     fmt_number(columns = c(`Adj R-Squared`, `Residual SE`, `P-value`),
                decimals = 4)

regh_model_table <- regh_summary_model %>%
  gt(rowname_col = " ",
     auto_align = FALSE) %>%
  tab_header(title = "(Figure-6) Model Statistics - Hound", 
             subtitle = "Regression of Average Pace on Heart Rate") %>%
     fmt_number(columns = c(`Adj R-Squared`, `Residual SE`, `P-value`),
                decimals = 4)

fit_reg_results <- data.frame(
  Runner = c("Collie", "Hound"),
  change_pace = c(coef(regc)["Avg.HR_Collie"], coef(regh)["Avg.HR_Hound"]),
  p_value_fr = c(summary(regc)$coefficients["Avg.HR_Collie","Pr(>|t|)"], summary(regh)$coefficients["Avg.HR_Hound","Pr(>|t|)"]),
  r_sqr_fr = c(summary(regc)$r.squared, summary(regh)$r.squared))
    
fitter_results <- fit_reg_results %>%
  gt() %>%
  tab_header(
    title = "(Figure-7) Runners Fitness",
    subtitle = "Average Pace on Heart Rate") %>%
    cols_label(change_pace = "Change in Pace", 
               p_value_fr = "P-value",
               r_sqr_fr = "R-squared") %>% 
  fmt_number(columns = c(change_pace, p_value_fr, r_sqr_fr), decimals = 4)

regc_model_table
regh_model_table

fitter_results

```

## Who improved their fitness more

To determine which runner improved their fitness more over time, we analyzed trends in both heart rate and pace by regressing each variable on date for each runner, as shown in Figure 8. For Collie, the regression coefficients were -0.0041 for heart rate and -0.0023 for pace. While the negative coefficient for heart rate indicates that Collie’s cardiovascular effort decreased over time (a clear sign of improved aerobic efficiency), the high p-value and r-squared for the average pace regression suggest that the observed decrease in pace over time is not statistically significant. In contrast, Hound’s coefficients were 0.0028 for heart rate and -0.0208 for pace, indicating that although his pace improved, it came alongside an increase in heart rate, reflecting pace decreases that required increased cardiovascular effort. Taken together, these results suggest that Collie improved their fitness more in terms of our definition of aerobic efficiency, as they were able to reduce heart rate over time while maintaining pace, whereas Hound’s improvements were in their pace alone and were accompanied by higher cardiovascular effort.

```{r, pace and HR trends over time}
#| echo: false
#| message: false
#| fig-align: center


regression_c_data <- reg_data %>% 
  dplyr::select(obs, Avg.HR, avg_pace_seconds, Date, Aerobic.TE, Runner) %>% 
  dplyr::filter(Runner == "Collie") %>%
  dplyr::filter(Aerobic.TE >= 3.5) %>%
  dplyr::select(-Runner) %>% 
  tidyr::drop_na()

reg1 <- stats::lm(Avg.HR ~ Date, data = regression_c_data)
reg2 <- stats::lm(avg_pace_seconds ~ Date, data = regression_c_data)

regression_h_data <- reg_data %>% 
  dplyr::select(obs, Avg.HR, Date, avg_pace_seconds, Aerobic.TE, Runner) %>% 
  dplyr::filter(Runner == "Hound") %>%
  dplyr::filter(Aerobic.TE >= 3.5) %>%
  dplyr::select(-Runner) %>% 
  tidyr::drop_na()

reg3 <- stats::lm(Avg.HR ~ Date, data = regression_h_data)
reg4 <- stats::lm(avg_pace_seconds ~ Date, data = regression_h_data) 

reg_results <- data.frame(
  Runner = c("Collie", "Hound"),
  change_hr = c(coef(reg1)["Date"], coef(reg3)["Date"]),
  p_value = c(summary(reg1)$coefficients["Date","Pr(>|t|)"], summary(reg3)$coefficients["Date","Pr(>|t|)"]),
  r_sqr = c(summary(reg1)$r.squared, summary(reg3)$r.squared),
  change_pace = c(coef(reg2)["Date"], coef(reg4)["Date"]),
  p_value = c(summary(reg2)$coefficients["Date","Pr(>|t|)"], summary(reg4)$coefficients["Date","Pr(>|t|)"]),
  r_sqr = c(summary(reg2)$r.squared, summary(reg4)$r.squared)
)
    
fitness_improvement <- reg_results %>%
  gt() %>%
  tab_header(
    title = "(Figure-8) Runners Fitness Improvement",
    subtitle = "Change in Pace and Heart Rate Over Time") %>%
    cols_label(change_hr = "Change in HR", 
               change_pace = "Change in Pace",
               p_value = "P-value",
               r_sqr = "R-squared",
               p_value.1 = "P-value",
               r_sqr.1 = "R-squared") %>% 
  fmt_number(columns = c(change_hr, change_pace, p_value, r_sqr, p_value.1, r_sqr.1), decimals = 4)

fitness_improvement
```

## Outdoor Running Performance

To determine Hound and Collie’s performance in outdoor runs we performed a seasonality analysis using average pace in seconds per kilometer recorded on a monthly basis for Collie in Figure-9 and Hound in Figure-10. We refined this data by taking the average of the data points in Figure-9 and Figure-10 to get a single data point for each month. We then performed a non-linear regression as seen in Figure-11 for Collie and Figure-12 for Hound. It is important to note that Hound had significantly more indoor runs that were removed when performing this seasonality analysis and after filtering, Collie only has one observation in December. When comparing Collie’s results in Figure-9 and Figure-11 and Hound’s results in Figure-10 and Figure-12 we observed a lower and more consistent average pace for Collie. This implies that outdoor running conditions on a month to month basis have less of an effect on Collie’s running performance than Hounds. Hounds' average pace follows a seasonality pattern that implies a slower pace in the colder months and a faster pace in the warmer months. This could be attributed to temperature or outdoor running conditions affecting Hound's pace. Hound’s February data includes two races in which Hound decreased their pace significantly which is most likely due to the run being in a competitive environment. When looking at Figure-9 it is apparent that Collie runs less in November and December than other months. As a coach we could encourage Collie to maintain their running volume year round to maintain consistency. Consistency is key as aerobic fitness can begin to decline in as little as 7-10 days [@2023Barbieri]. Overall, Collie’s current level of consistency is a positive from a coaching perspective as we know that outdoor conditions will have a limited affect on their running ability and effort and extending Collie’s consistency through all months of the year is more feasible than trying to build a new habit of consistency for Hound.

```{r, seasonality}
#| echo: false
#| message: false
#| fig-align: center

season_dat_collie <- run_data %>%
  dplyr::filter(Runner == "Collie") %>% 
  dplyr::filter(!grepl("Track|Tread", Activity.Type))%>%
  dplyr::filter(avg_pace_seconds<="380") %>% 
  dplyr::select(Date, avg_pace_seconds, Runner, Activity.Type )%>%
  tsibble::as_tsibble(key = Runner, index = Date) %>%
  tsibble::group_by_key() %>% 
  tsibble::index_by(freq = ~ yearmonth(.)) %>% 
  stats::na.omit()
 


graph_collie <- season_dat_collie %>%
  tidyr::separate(freq, 
                  into = c("Year", "Month"), 
                  sep = " ", 
                  convert = TRUE, 
                  remove = FALSE) %>% 
   dplyr::mutate(Month = factor(Month, levels = c("Jan", "Feb", "Mar", "Apr","May" ,"Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))) %>% 
  dplyr::group_by(Month) %>% 
  dplyr::mutate(avg_mon_pace = mean(avg_pace_seconds),.groups = "drop")
  

season_dat_Hound <- run_data %>%
  dplyr::filter(Runner == "Hound") %>% 
  dplyr::filter(!grepl("Track|Tread", Activity.Type))%>%
  dplyr::filter(avg_pace_seconds>="300") %>% 
  dplyr::filter(!duplicated(Date)) %>% 
  dplyr::select(Date, avg_pace_seconds, Runner, Activity.Type ) %>% 
  tsibble::as_tsibble(key = Runner, index = Date) %>%
  tsibble::group_by_key() %>% 
  tsibble::index_by(freq = ~ yearmonth(.)) %>%
  stats::na.omit() 

graph_hound <- season_dat_Hound %>%
  tidyr::separate(freq, 
                  into = c("Year", "Month"), 
                  sep = " ", 
                  convert = TRUE, 
                  remove = FALSE) %>% 
  dplyr::mutate(Month = factor(Month, levels = c("Jan", "Feb", "Mar", "Apr","May" ,"Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"))) %>% 
  dplyr::group_by(Month) %>% 
  dplyr::mutate(avg_mon_pace = mean(avg_pace_seconds),.groups = "drop")

  

plot_c <- graph_collie %>% 
  ggplot2::ggplot(ggplot2::aes(x = Month, y = avg_pace_seconds)) + 
  geom_point(col= "darkgreen") +
ggplot2::labs(title = "(Figure-9) Average Pace by Month",
                subtitle = "Seasonality Analysis Collie",
                y = "Avg Pace (Seconds)")
  
plot_h <- graph_hound %>% 
  ggplot2::ggplot(ggplot2::aes(x = Month, y = avg_pace_seconds)) + 
  geom_point(col = "purple") +
ggplot2::labs(title = "(Figure-10) Average Pace by Month",
                subtitle = "Seasonality Analysis Hound",
                y = "Avg Pace (Seconds)")
  


plot_c /plot_h
```

```{r, non-linear reggression on seasonality results}
#| echo: false
#| message: false
#| fig-align: center

#Numbering months to allow geom_smooth function to work. 
graph_collie <- graph_collie %>% 
  dplyr::mutate(Month_num = month(Date , label = FALSE)) 

graph_hound <- graph_hound %>% 
  dplyr::mutate(Month_num = month(Date , label = FALSE))

plot_c_nonlinear <- graph_collie %>% 
  ggplot2::ggplot(aes(x = Month_num, y = avg_mon_pace)) + 
  geom_point() + 
  geom_smooth( method = "lm", formula = y ~ splines::bs(x, knots = c(4, 9), degree = 3), se = FALSE , col = "darkgreen") + 
  scale_x_continuous(n.breaks = 12) +
  ggplot2::labs(
    title = "(Figure-11) Average Pace By Month",
    subtitle = "Seasonality Analysis Collie",
    x = "Month (Jan-Dec)",
    y = "Avg Pace"
  )

plot_h_nonlinear <- graph_hound %>% 
  ggplot2::ggplot(aes(x = Month_num, y = avg_mon_pace)) +
  geom_point() + 
  geom_smooth( method = "lm", formula = y ~ splines::bs(x, knots = c(2, 6, 8), degree = 3), se = FALSE, col = "purple" ) + 
  scale_x_continuous(n.breaks = 12) +
  ggplot2::labs(
    title = "(Figure-12) Average Pace By Month",
    subtitle = "Seasonality Analysis Hound",
    x = "Month (Jan-Dec)",
    y = "Avg Pace"
  )

plot_c_nonlinear/plot_h_nonlinear

```

## Who is Better to Coach

One unique value we are exposed to with this data set is each runner's Aerobic TE. This measurement is scored on a 0-5 scale and uses many individualized, along with run-specific variables, to determine how effective the run is on their fitness. In this analysis, we considered Aerobic TE to be a strong measure for understanding which runner exerts more effort. Running at a higher Aerobic TE is more difficult to maintain and requires discipline to overcome the strain that running has on an individual's body. Figure 14 depicts a histogram of each runners Aerobic TE from their available runs. This data was  filtering in the same way as we did when determining who is fitter, removing any observations that were seen to be not run specific. When coaching, having an athlete who is more dedicated to improvement and willing to exert maximum effort will typically see results that follow sooner. In the case of the coach, higher effort not only benefits the athletes' improvement in running but can also help the coach understand if their techniques or trials are working by quantifying results faster, compared to coaching someone where improvement may be slower due to a lack of maximum effort. Figure 14 provides a depiction of which runner is exerting more effort more often, as well as a  red line that indicates the mean Aerobic TE for each runner over all the analyzed runs. It is clear that Collie maintains higher Aerobic TE where they are observed running in the “Highly improving”[@Garmin] area compared to Hound, who is generally seen to be running in the “improving” area. It is also important to note that Collie has many observations at an Aerobic TE of 5, which is considered “Overreaching” and can potentially be considered dangerous; however, coaching someone to exert less effort is much easier than convincing someone to exert more. In our evaluation of which runner to coach, we determined that Collie would be a better fit and will likely see a greater impact from additional support and professional guidance.

```{r}
#| echo: false
#| message: false
#| warning: false
#| fig-align: center

regc_dat$Aerobic.TE <- as.numeric(regc_dat$Aerobic.TE)
regh_dat$Aerobic.TE <- as.numeric(regh_dat$Aerobic.TE)

te_analysis_c <- regc_dat %>% 
  dplyr::select(Aerobic.TE) 

te_analysis_h <- regh_dat %>% 
  dplyr::select(Aerobic.TE) 


c_hist <- te_analysis_c %>% 
  ggplot2::ggplot(ggplot2::aes(x = Aerobic.TE, )) +
  ggplot2::geom_histogram(aes(y = after_stat(width*density)) ,
                          bins = 40, col = "black", fill = "darkgreen")+ 
   ggplot2::geom_vline(ggplot2::aes(xintercept = mean(Aerobic.TE),col = Runner), col = "red")+
  scale_y_continuous(labels = scales::percent_format())+
  ggplot2::scale_x_continuous(limits =c(1,5.1)) +
  ggplot2::labs(
    title = "(Figure-14) Distribution of Runs by Aerobic.TE",
    subtitle = "Collie",
    y = "Percentage of Observations",
    x = "Aerobic.TE Score"
  )

h_hist <- te_analysis_h %>% 
  ggplot2::ggplot(ggplot2::aes(x = Aerobic.TE, )) +
  ggplot2::geom_histogram(aes(y = after_stat(width*density)) , 
                          bins = 40, col = "black", fill = "purple")+ 
   ggplot2::geom_vline(ggplot2::aes(xintercept = mean(Aerobic.TE),col = Runner), col = "red")+
  scale_y_continuous(labels = scales::percent_format())+
  ggplot2::scale_x_continuous(limits =c(1,5.1)) +
  ggplot2::labs(
    title = "(Figure-15) Distribution of Runs by Aerobic.TE",
    subtitle = "Hound",
    y = "Percentage of Observations",
    x = "Aerobic.TE Score"
  )
#[@RunningEfficiency]
c_hist/h_hist
```
